Beta Diversity through Elevation

### 2021-09-22 code/ms1_beta_diversity.Rmd
analysis reboot

### Prepare environment
```{r message=FALSE, warning=FALSE}
library(sf)
library(gdm)
library(patchwork)
library(bipartite)
library(vegan)
library(ecodist)
library(gt)
library(tidyverse)
library(lubridate)
```

### Define functions
```{r}
# A function for converting a geometry column back to x and y columns
# https://github.com/r-spatial/sf/issues/231
sfc_as_cols <- function(x, names = c("x","y")) {
  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
  ret <- sf::st_coordinates(x)
  ret <- tibble::as_tibble(ret)
  stopifnot(length(names) == ncol(ret))
  x <- x[ , !names(x) %in% names]
  ret <- setNames(ret,names)
  dplyr::bind_cols(x,ret)
}

# Function to convert networks to webs arrays
web_prep <- function(net) {

  out <- net %>%
    filter(caste != "male") %>%
    group_by(site, bb.sp, plant.sp) %>%
    summarize(freq = n()) %>%
    group_by(site) %>%
    mutate(bb.sp.count = length(unique(bb.sp)),
           plant.sp.count = length(unique(plant.sp))) %>%
    ungroup() %>%
    dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs(type.out = "array")
}

# Function for betalinkr settings and conversion to tibble()
betalinkr_set1 <- function(x) {
  betalinkr_multi(x, partition.st = TRUE, partition.rr = TRUE, # partition both species turnover and richness/replacement
                  binary = TRUE, # treating beta-diversity as binary is an appropriate complement to our abundance analysis
                  index = "jaccard") %>% # the Jaccard index is best, I think, for binary analysis
  as_tibble() # for convenience
}

# Function for converting visitation data into site-pair table for GDM analysis

interaction_gdm_prep <- function(x) {

@@

  # Prep site-pairs for GDM analysis
  sitepair <- formatsitepair(spdat, 
                             bioFormat = 1, # i.e. samples as rows, species as columns, site-dates as rows
                             siteColumn = "site", 
                             dist = "jaccard", # I think Jaccard is the right choice
                             XColumn = "x", YColumn = "y", # spatial columns
                             predData = envdat, 
                             abundance = FALSE) %>% # abundance = FALSE means binary
    filter(s1.elev.mean != s2.elev.mean) # this drops within-site comparisons
}

# Function for preparing the floral survey biodata input to gdm's formatsitepair function
biodata_prep_survey <- function(surv) {
  surv %>%
    filter(flower.cover > 0) %>%
    mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
    group_by(site, plant.sp) %>%
    summarize(cover = sum(flower.cover)) %>%
    mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
    spread(plant.sp, cover) %>%
    replace(is.na(.), 0) %>%
    arrange(site)
}

# Function for preparing the floral survey predictor data input to gdm's formatsitepair function
predData_prep_survey <- function(surv) {
  surv %>%
    filter(flower.cover > 0) %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    mutate(mowing = if_else(management == "mowing", 1, 0),
           grasing = if_else(management == "grasing", 1, 0),
           none = if_else(management == "none", 1, 0)) %>%
    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
    distinct() %>%
    arrange(site) 
}

# Function for preparing the visitation biodata input to gdm's formatsitepair function
biodata_prep_BB <- function(net) {
  net %>%
    filter(caste != "male") %>%
    group_by(site, bb.sp) %>%
    summarize(count = n()) %>%
    mutate(count = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
    spread(bb.sp, count) %>%
    replace(is.na(.), 0) %>%
    ungroup()
}

# Function for preparing the visitation predictor data input to gdm's formatsitepair function
predData_prep_BB <- function(net) {
  net %>%
    filter(caste != "male") %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    mutate(mowing = if_else(management == "mowing", 1, 0),
           grasing = if_else(management == "grasing", 1, 0),
           none = if_else(management == "none", 1, 0)) %>%
    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
    unique() %>%
    arrange(site) 
}

# Function with setting for gdm's formatsitepair function
format_set <- function(bio, pred) {
  formatsitepair(bio, bioFormat = 1, siteColumn = "site", 
                 dist = "jaccard", XColumn = "x", YColumn = "y", 
                 predData = pred, abundance = FALSE) %>%
  filter(s1.elev.mean != s2.elev.mean) # this avoids comparing repeated measures of the same site (taking advantage of the fact that each site has a unique elevation in our dataset)
}

# Function for tabulating gdm.varImp output 
gdm_tab <- function(x, response) {
  x %>%
    as.data.frame() %>%
    rownames_to_column(var = "predictor") %>%
    mutate(response = rep(response, n()),
           Predictor = case_when(
             predictor == "elev.mean" ~ "elevation",
             predictor == "samples" ~ "sampling freq.",
             predictor == "mowing" ~ "mowing",
             predictor == "grasing" ~ "grasing",
             predictor == "none" ~ "none")) %>%
    select(response, Predictor, Importance, pval) %>%
    tibble()
}
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% #
  mutate(site = factor(site),
         tree_line = factor(tree_line)) %>% 
  semi_join(read_csv("../data/processed_data/network.csv"), by = "site") %>% # drop sites that were not sampled
  st_as_sf(coords = c("lon", "lat")) %>%
  st_set_crs(4326) %>%
  st_transform(5684) %>%
  sfc_as_cols() %>%
  as_tibble() %>%
  dplyr::select(-geometry)

# Write to csv for archiving
write_csv(site_data, "../Dryad_ms1/site_data.csv")
```

### Load network and survey data
```{r}
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste  != "male") %>%
  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>% 
  mutate(year = factor(year))

@@

# Write to csv for archiving
write_csv(net, "../Dryad_ms1//visitation_data.csv")
write_csv(survey, "../Dryad_ms1//floral_survey_data.csv")
```

### Get beta-diversity for BBs and plants (not their interactions)
```{r}
# Bumble bee beta-diversity based on interaction sampling
bb_beta <- net %>%
  group_by(site, bb.sp) %>%
  summarize(count = n()) %>% 
  pivot_wider(names_from = bb.sp, values_from = count) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("site") %>%
  vegdist(method = "jaccard", binary = TRUE) %>% # Jaccard and binary settings make this comparable to the way I'm running betalinkr
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "BB")

@@

# Join into single data frame for visualization
species_beta <- full_join(bb_beta, fl_beta) %>%
  left_join(site_data, by = c("site1" = "site")) %>% # add site data
  dplyr::select(site1, site2, FL, BB, # first for site 1
                elev1 = elev.mean, x1 = x, y1 = y, mgmt1 = management) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, FL, BB, # then for site 2
                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y, mgmt1, mgmt2 = management) %>%
  mutate(site1 = factor(site1), 
         site2 = factor(site2),
         elev.diff = elev1 - elev2,
         mean.elev = (elev1 + elev2)/2, mgmt.diff = if_else(mgmt1 == mgmt2, 0, 1)) %>%
  pivot_longer(cols = c(BB, FL), names_to = "group", values_to = "beta") %>%
  filter(elev.diff >= 0) # For each site pair, one will be higher and the other lower; thus, when their order is reversed only the sign of the difference changes. Since no two sites have exactly the same elevation, this deduplicates the data by dropping the ordering of a given pair of sites (one of the two orderings) that yields a negative elevation difference while retaining the alternative ordering that yields a positive elevation difference. By >= 0, I retain comparisons of a site to itself. I drop (in the ggplot call) within-site comparisons when I plot against elevation difference, and I isolate within-site comparison when I plot against time. In the latter case, it is also necessary to deduplicate, since each pair of site dates occurs in A-B and B-A. For that, I filter (again, in the ggplot call) so that yday.diff > 0, similarly dropping the orderings that yield negative differences. Note that all this deduplication nonsense is not necessary for the betalinkr call because it deduplicates internally.

# fl_beta_dist <- survey %>%
#   group_by(site, plant.sp) %>%
#   summarize(count = sum(flower.cover)) %>%
#   pivot_wider(names_from = plant.sp, values_from = count) %>%
#   replace(is.na(.), 0) %>%
#   column_to_rownames("site") %>%
#   vegdist(method = "jaccard", binary = TRUE, upper = TRUE) 
# 
# bb_beta_dist <- net %>%
#   group_by(site, bb.sp) %>%
#   summarize(count = n()) %>% 
#   pivot_wider(names_from = bb.sp, values_from = count) %>%
#   replace(is.na(.), 0) %>%
#   column_to_rownames("site") %>%
#   vegdist(method = "jaccard", binary = TRUE)
```

### Tabulate webs
S	= beta_S, the dissimilarity in species composition
OS = beta_OS, the dissimilarity (component) explained by "rewiring" among shared species
WN = beta_WN, the dissimilarity between the two networks
ST = beta_ST, the dissimilarity (component) explained by difference in species community composition 
```{r message=FALSE}
# Create webs
net_web <- web_prep(net)

@@

# Write to csv for archiving
write_csv(beta_mit_data, "../Dryad_ms1/partitioned_beta_diversity")
```

### Plot interaction beta-diversity against elevation difference
```{r}
# Set palette
p1 <- c("OS" = "#e7298a", "ST" = "#66a61e", "WN" = "black")

>>>>

# 1. Prepare environment
## 1.1 Packages 
```{r message=FALSE, warning=FALSE}
library(sf)
library(magick)
library(gdm)
library(patchwork)
library(bipartite)
library(vegan)
library(ecodist)
library(flextable)
library(tidyverse)
library(lubridate)
library(cowplot)
```
## 1.2 Define functions
### 1.2.1 `sfc_as_cols`: convert geometry column back to x and y columns
```{r}
# A function for converting a geometry column back to x and y columns
# https://github.com/r-spatial/sf/issues/231
sfc_as_cols <- function(x, names = c("x","y")) {
  stopifnot(inherits(x,"sf") && inherits(sf::st_geometry(x),"sfc_POINT"))
  ret <- sf::st_coordinates(x)
  ret <- tibble::as_tibble(ret)
  stopifnot(length(names) == ncol(ret))
  x <- x[ , !names(x) %in% names]
  ret <- setNames(ret,names)
  dplyr::bind_cols(x,ret)
}
```

### 1.2.2 `web_prep`: convert networks to webs arrays
```{r}
# Function to convert networks to webs arrays
web_prep <- function(net) {

  out <- net %>%
    filter(caste != "male") %>%
    group_by(site, bb.sp, plant.sp) %>%
    summarize(freq = n()) %>%
    group_by(site) %>%
    mutate(bb.sp.count = length(unique(bb.sp)),
           plant.sp.count = length(unique(plant.sp))) %>%
    ungroup() %>%
    dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
    data.frame() %>% # convert to data frame
    frame2webs(type.out = "array")
}
```

### 1.2.3 `betalinkr_set1`: betalinkr settings and conversion to tibble
```{r}
# Function for betalinkr settings and conversion to tibble()
betalinkr_set1 <- function(x) {
  betalinkr_multi(x, partition.st = TRUE, partition.rr = TRUE, # partition both species turnover and richness/replacement
                  binary = TRUE, # treating beta-diversity as binary is an appropriate complement to our abundance analysis
                  index = "jaccard") %>% # the Jaccard index is best, I think, for binary analysis
  as_tibble() # for convenience
}
```

### 1.2.4 `interaction_gdm_prep`: convert visitation data into site-pair table for GDM analysis
```{r}
# Function for converting visitation data into site-pair table for GDM analysis 
interaction_gdm_prep <- function(x) {

@@

  # Prep site-pairs for GDM analysis
  sitepair <- formatsitepair(spdat, 
                             bioFormat = 1, # i.e. samples as rows, species as columns, site-dates as rows
                             siteColumn = "site", 
                             dist = "jaccard", # I think Jaccard is the right choice
                             XColumn = "x", YColumn = "y", # spatial columns
                             predData = envdat, 
                             abundance = FALSE) %>% # abundance = FALSE means binary
    filter(s1.elev.mean != s2.elev.mean) # this drops within-site comparisons
}
```

### 1.2.5 `biodata_prep_survey`: prepare the floral survey biodata input to gdm's `formatsitepair()` function
```{r}
# Function for preparing the floral survey biodata input to gdm's formatsitepair function
biodata_prep_survey <- function(surv) {
  surv %>%
    filter(flower.cover > 0) %>%
    mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
    group_by(site, plant.sp) %>%
    summarize(cover = sum(flower.cover)) %>%
    mutate(cover = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
    spread(plant.sp, cover) %>%
    replace(is.na(.), 0) %>%
    arrange(site)
}
```

### 1.2.6 `predData_prep_survey`: prepare the floral survey predictor data input to gdm's `formatsitepair()` function
```{r}
# Function for preparing the floral survey predictor data input to gdm's formatsitepair function
predData_prep_survey <- function(surv) {
  surv %>%
    filter(flower.cover > 0) %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    mutate(mowing = if_else(management == "mowing", 1, 0),
           grasing = if_else(management == "grasing", 1, 0),
           none = if_else(management == "none", 1, 0)) %>%
    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
    distinct() %>%
    arrange(site) 
}
```

### 1.2.7 `biodata_prep_BB`: prepare the visitation biodata input to gdm's `formatsitepair()` function
```{r}
# Function for preparing the visitation biodata input to gdm's formatsitepair function
biodata_prep_BB <- function(net) {
  net %>%
    filter(caste != "male") %>%
    group_by(site, bb.sp) %>%
    summarize(count = n()) %>%
    mutate(count = rep(1, n())) %>% # this makes the data binary, which I think is what I want here; otherwise, just comment out
    spread(bb.sp, count) %>%
    replace(is.na(.), 0) %>%
    ungroup()
}
```

### 1.2.8 `predData_prep_BB`: prepare the visitation predictor data input to gdm's `formatsitepair()` function
```{r}
# Function for preparing the visitation predictor data input to gdm's formatsitepair function
predData_prep_BB <- function(net) {
  net %>%
    filter(caste != "male") %>%
    group_by(site) %>%
    summarize(samples = length(unique(date))) %>%
    dplyr::select(site, samples) %>%
    left_join(site_data) %>%
    ungroup() %>%
    mutate(mowing = if_else(management == "mowing", 1, 0),
           grasing = if_else(management == "grasing", 1, 0),
           none = if_else(management == "none", 1, 0)) %>%
    dplyr::select(site, elev.mean, x, y, samples, mowing, grasing, none) %>%
    unique() %>%
    arrange(site) 
}
```

### 1.2.9 `format_set`: setting for gdm's `formatsitepair()` function
```{r}
# Function with setting for gdm's formatsitepair function
format_set <- function(bio, pred) {
  formatsitepair(bio, bioFormat = 1, siteColumn = "site", 
                 dist = "jaccard", XColumn = "x", YColumn = "y", 
                 predData = pred, abundance = FALSE) %>%
  filter(s1.elev.mean != s2.elev.mean) # this avoids comparing repeated measures of the same site (taking advantage of the fact that each site has a unique elevation in our dataset)
} 
```

### 1.2.10 `gdm_tab`: tabulate `gdm.varImp()` output 
```{r}
# Function for tabulating gdm.varImp output 
gdm_tab <- function(x, response) {
  x %>%
    as.data.frame() %>%
    rownames_to_column(var = "predictor") %>%
    mutate(model = rep(response, n()),
           predictor = case_when(
             predictor == "elev.mean" ~ "elevation",
             predictor == "samples" ~ "samples",
             predictor == "Geographic" ~ "distance")
             # predictor == "mowing" ~ "mowing",
             # predictor == "grasing" ~ "grazing",
             # predictor == "none" ~ "none")
           ) %>%
    mutate(importance = round(importance, 2),
           model.var.exp = round(model.var.exp, 2)) %>%
    select(Model = model, Variable = predictor, 
           "Importance [%]" = importance, Var.pval = pval, 
           Model.pval = model.pval, "Model.dev.exp [%]" = model.var.exp) %>%
    tibble() %>%
    mutate(Variable = factor(Variable, levels = c("elevation", "distance", "samples"))) %>%
    arrange(Variable)
}
```

### 1.2.11 `isplineExtract2`: Fixes a bug in `isplineExtract()` that causes it to break when non-default number of splines are used
```{r}
##Extracts Ispline data from a gdm model
isplineExtract2 <- function (model){
  ###########################
  #model = gdmOb
  ###########################
  ##error checking
  ##checks to make sure a gdm model is given
  if(!is(model, "gdm")){
    stop("model argument must be of class = 'gdm'.")
  }

  ##Collects or sets simple data
  options(warn.FPU = FALSE)
  PSAMPLE <- 200
  preddata <- rep(0, times = PSAMPLE)
  pn <- model$predictors
  nPreds <- length(pn)
  yDat <- xDat <- matrix(0,PSAMPLE,nPreds)
  colnames(yDat) <- colnames(xDat) <- pn
  pmin <- 1
  pmax <- PSAMPLE

  ##cycles through each prodictor and fills the spline matrices
  splineindex <- 1
  for (i in 1:nPreds){ 
    #i<-1
    numsplines <- model$splines[i]
    z <- .C("GetPredictorPlotData", 
            pdata = as.double(preddata), 
            as.integer(PSAMPLE), 
            as.double(model$coefficients[splineindex:(splineindex + numsplines - 1)]), 
            as.double(model$knots[splineindex:(splineindex + numsplines - 1)]), 
            as.integer(numsplines),
            PACKAGE = "gdm")
    yDat[,i] <- z$pdata
    pmin <- pmin + PSAMPLE
    pmax <- pmax + PSAMPLE
    xDat[,i] <-  seq(from=model$knots[[(i*numsplines)-(numsplines-1)]],to=model$knots[[(i*numsplines)]], length=PSAMPLE)
    splineindex <- splineindex + numsplines
  }

  ##lists and returns matrices
  outData <- list(x=xDat,y=yDat)
  return(outData)
}
```

### 1.2.12 `plotUncertainty2` Tweaked to call `isplineExtract2` instead of the broken version
```{r}
##########################################################################
plotUncertainty2 <- function(spTable, sampleSites, bsIters, geo=FALSE, splines=NULL, 
                           knots=NULL, splineCol="blue", errCol="grey80", 
                           plot.linewidth=2.0, plot.layout=c(2,2), parallel=FALSE,
                           cores=2){
  ##A function to plot the uncertantiy of each variable from a GDM model
  #################
  #spTable <- sitePairTab          ##the input site-pair table to subsample from
  #sampleSites <- 0.9     ##fraction of sites that should be retained from site pair table 
  #bsIters <- 50       ##the number of times the site-pair table should be sampled
  #geo <- F              ##rather or not the gdm model takes geography into account, see gdm
  #splines <- NULL       ##splines gdm setting, see gdm
  #knots <- NULL         ##knots gdm setting, see gdm
  #splineCol <- "blue"    ##color of the center line
  #errCol <- "grey80"        ##color of the uncertainty polygon
  #plot.linewidth <- 2.0    ##line width of the center line
  #plot.layout <- c(3,3)    ##number of plots per page
  #parallel <- F            ##rather or not the sampling should happen in parallel processing, to speed it up
  #cores <- 6               ##number of cores to if parallel processing
  #################
  ##function breaks and warnings
  ##makes sure that table is a properly formated site-pair table
  if(!is(spTable, "gdmData")){
    warning("The spTable object is not of class 'gdmData'. See the formatsitepair function for help.")
  }
  ##checks to makes sure table is a matrix or table frame
  if(!(is(spTable, "gdmData") | is(spTable, "matrix") | is(spTable, "data.frame"))){
    stop("The spTable object must be of class 'gdmData', 'matrix', or 'data.frame'.")
  }

  ##sanity check on the data table
  if(ncol(spTable) < 6){
    stop("spTable object requires at least 6 columns: Observed, weights, s1.xCoord, s1.yCoord, s2.xCoord, s2.yCoord")
  }  
  if(nrow(spTable) < 1){
    stop("spTable object has < 1 rows.")
  }

  ##checks that geo has either TRUE or FALSE
  if(!(geo==TRUE | geo==FALSE)){
    stop("geo argument must be either TRUE or FALSE")
  }
  ##makes sure splines is a numeric vector
  if(is.null(splines)==FALSE & !is(splines, "numeric")){
    stop("splines object must of of class = 'numeric'.")
  }
  ##checks knots inputs
  if(is.null(knots)==FALSE & !is(knots, "numeric")){
    stop("knots object must of of class = 'numeric'.")
  }

  ##checks that parallel has either TRUE or FALSE
  if(!(parallel==TRUE | parallel==FALSE)){
    stop("parallel argument must be either TRUE or FALSE")
  }
  ##makes sure that cores has a value when parallel is true
  if(parallel==TRUE & is.null(cores)==TRUE){
    stop("If parallel==TRUE, the number of cores must be specified")
  }
  ##makes sure that cores is a positive integer 
  if((is.null(cores)==FALSE & is.numeric(cores)==FALSE) | cores<1){
    stop("argument cores needs to be a positive integer")
  }

  ##makes sure that bsIters is a positive integer 
  if((is.null(bsIters)==FALSE & is.numeric(bsIters)==FALSE) | bsIters<1){
    stop("argument bsIters needs to be a positive integer")
  }
  ##makes sure sampleSites is a number
  if(is.numeric(sampleSites)==FALSE){
    stop("sampleSites must be a number between 0 and 1")
  }
  ##makes sure that sampleSites is between 0 and 1
  if(sampleSites < 0){
    stop("sampleSites must be a number between 0 and 1")
  }
  if(sampleSites > 1){
    stop("sampleSites must be a number between 0 and 1")
  }
  if(sampleSites==0){
    stop("a sampleSites value of 0 will remove all sites from the analysis (bad).")
  }
  ##double makes sure these values are integers, seems to truncate if not
  cores <- as.integer(cores)
  bsIters <- as.integer(bsIters)

  ##assign k to prevent issues to cran checking
  k <- NULL

  ##makes copies of the site-pair table in order to randomly subsample each one differently
  lstSP <- lapply(1:bsIters, function(i){spTable})

  ##runs parallel if desired by the users
  if(parallel==TRUE){
    ##loads libraries
    #require(foreach)
    #require(doParallel)
    #requireNamespace("foreach")
    #requireNamespace("parallel")
    #requireNamespace("doParallel")

    ##sets cores
    cl <- makeCluster(cores, outfile="")
    registerDoParallel(cl)
    ##first removes a number of sites according to input
    subSamps <- foreach(k=1:length(lstSP), .verbose=F, .packages=c("gdm")) %dopar%
                      removeSitesFromSitePair(lstSP[[k]], sampleSites=sampleSites)
    ##models the subsamples
    gdmMods <- foreach(k=1:length(subSamps), .verbose=F, .packages=c("gdm")) %dopar%
                     #gdmMods <- try(foreach(k=1, .verbose=F, .packages=c("gdm")) %dopar%
                     gdm(subSamps[[k]], geo=geo, splines=splines, knots=knots)
    stopCluster(cl)
  }else{
    ##first removes a number of sites according to input
    subSamps <- lapply(lstSP, removeSitesFromSitePair, sampleSites=sampleSites)
    ##models the subsamples
    gdmMods <- lapply(subSamps, gdm, geo=geo, splines=splines, knots=knots)
  }

  ##models the full gdm
  fullGDMmodel <- gdm(spTable, geo=geo, splines=splines, knots=knots)

  ##Extracts the splines for each model
  exUncertSplines <- lapply(gdmMods, isplineExtract2)
  fullGDMsplines <- isplineExtract2(fullGDMmodel)

  ##get the names of the predictor variables
  predVars <- colnames(exUncertSplines[[1]][[1]])

  ##establish what plot layout to use
  thisplot <- 0
  one_page_per_plot <- FALSE
  if ((plot.layout[1]==1) && (plot.layout[2]==1)){
    one_page_per_plot <- TRUE
  }else{
    par(mfrow=plot.layout)
  }

  ##sets the plotting minimum and maximum y-values 
  totalYmin <- Inf
  totalYmax <- -Inf

  ##determines the bounds of the plots
  for(p in 1:length(predVars)){
    #p=1
    predV <- predVars[p]

    ##gets the minimum and maximum, to set the ploting extent
    for(nm in 1:length(exUncertSplines)){
      #nm=1
      selPlot <- exUncertSplines[[nm]]

      spYmax <- max(selPlot[[2]][,predV])
      spYmin <- min(selPlot[[2]][,predV])

      totalYmax <- max(c(totalYmax, spYmax))
      totalYmin <- min(c(totalYmin, spYmin))
    }
  }

  ##plots by variable
  for(p in 1:length(predVars)){
    #p <- 1
    predV <- predVars[p]

    ##sets the plotting minimum and maximum x-values
    totalXmin <- Inf
    totalXmax <- -Inf
    ##gets the minimum and maximum, to set the ploting extent
    for(nm in 1:length(exUncertSplines)){
      #nm=1
      selPlot <- exUncertSplines[[nm]]

      spXmax <- max(selPlot[[1]][,predV])
      spXmin <- min(selPlot[[1]][,predV])

      if(spXmax > totalXmax){totalXmax = spXmax}
      if(spXmin < totalXmin){totalXmin = spXmin}
    }

    ##checks to make sure that there is some significance, if there is the the data is plotted
    if(totalYmax!=0){
      ##add mean of subsets to plot
      plotX <- NULL
      plotY <- NULL
      byVarMatX <- NULL
      byVarMatY <- NULL
      ##create matrices based on the variable and its x and y location for each model iteration
      for(nn in 1:length(exUncertSplines)){
        #nn=1
        plotX[[nn]] <- exUncertSplines[[nn]][[1]]
        plotY[[nn]] <- exUncertSplines[[nn]][[2]]
        byVarMatY <- cbind(byVarMatY, plotY[[nn]][,predV])
        byVarMatX <- cbind(byVarMatX, plotX[[nn]][,predV])
      }
      ##gets spline data from the full gdm model
      fullPlotX <- fullGDMsplines[[1]]
      fullPlotX <- fullPlotX[,predV]
      fullPlotY <- fullGDMsplines[[2]]
      fullPlotY <- fullPlotY[,predV]

      ##calculates the confidence intervals for plotting
      sdX <- apply(as.matrix(byVarMatX), 1, sd)
      sdY <- apply(as.matrix(byVarMatY), 1, sd)
      highBoundX <- fullPlotX + sdX
      lowBoundX <- fullPlotX - sdX
      highBoundY <- fullPlotY + sdY
      lowBoundY <- fullPlotY - sdY

      ##collects the data to be used in the rug plot
      if(predV=="Geographic"){
        ##calculates unique eucildian distance between sites
        rugData <- unique(sqrt(((spTable$s1.xCoord-spTable$s2.xCoord)^2)+((spTable$s1.yCoord-spTable$s2.yCoord)^2)))
      }else{
        ##gets unique values of variable data
        varDat <- grep(predV, colnames(spTable))
        rugData <- unique(c(spTable[,c(varDat[1])], spTable[,c(varDat[2])]))
      }

      ##plots one graph per page, unless specified otherwise
      if (one_page_per_plot){
        dev.new()
        dev.next()
      }else{
        thisplot <- thisplot + 1
        if(thisplot > (plot.layout[1] * plot.layout[2])){    			
          thisplot <- 1
          par(mfrow=plot.layout)	
        }
      }
      #settings <- par(pars)
      ##plots mean data line and polygon of uncertanty 
      plot(NULL, xlim=c(totalXmin, totalXmax), ylim=c(totalYmin, totalYmax), xlab=predV, ylab="Partial Ecological Distance")
      polygon(c(lowBoundX, rev(highBoundX)), c(lowBoundY, rev(highBoundY)), col=errCol, border=NA)
      lines(fullPlotX, fullPlotY, col=splineCol, lwd=plot.linewidth)
      rug(rugData)
    }
  }
}
##########################################################################
```

## 1.3 Load data
```{r}
net <- read_csv("../data/processed_data/network.csv") %>%
  filter(caste  != "male") %>%
  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>% 
  mutate(year = factor(year))

@@

site_data <- read_csv("../data/processed_data/site_data.csv") %>%
  dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean, elev.class, elev.class2)) %>% #
  mutate(site = factor(site),
         tree_line = factor(tree_line)) %>% 
  semi_join(net, by = "site") %>% # drop sites that were not sampled
  st_as_sf(coords = c("lon", "lat")) %>%
  st_set_crs(4326) %>%
  st_transform(5684) %>%
  sfc_as_cols() %>%
  as_tibble() %>%
  dplyr::select(-geometry)
```

## 1.4 Load images
```{r}
BB <- image_read("../ms_1/figures/phylopic_BB.png")
FL <- image_read("../ms_1/figures/phylopic_FL.png")
BBxFL <- image_read("../ms_1/figures/phylopic_BBxFL.png")
```

# 2. Process data
## 2.1 BB and FL beta-diversity
```{r}
# Bumble bee beta-diversity based on interaction sampling
bb_beta <- net %>%
  group_by(site, bb.sp) %>%
  summarize(count = n()) %>% 
  pivot_wider(names_from = bb.sp, values_from = count) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("site") %>%
  vegdist(method = "jaccard", binary = TRUE) %>% # Jaccard and binary settings make this comparable to the way I'm running betalinkr
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "BB")

@@

# Join into single data frame for visualization
species_beta <- full_join(bb_beta, fl_beta) %>%
  left_join(site_data, by = c("site1" = "site")) %>% # add site data
  dplyr::select(site1, site2, FL, BB, # first for site 1
                elev1 = elev.mean, x1 = x, y1 = y, mgmt1 = management) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, FL, BB, # then for site 2
                elev1, elev2 = elev.mean, x1, y1, x2 = x, y2 = y, mgmt1, mgmt2 = management) %>%
  mutate(site1 = factor(site1), 
         site2 = factor(site2),
         elev.diff = elev1 - elev2,
         mean.elev = (elev1 + elev2)/2, mgmt.diff = if_else(mgmt1 == mgmt2, 0, 1)) %>%
  pivot_longer(cols = c(BB, FL), names_to = "group", values_to = "beta") %>%
  filter(elev.diff >= 0)

# For each site pair, one will be higher and the other lower; thus, when their order is reversed only the sign of the difference changes. Since no two sites have exactly the same elevation, `filter(elev.diff >= 0)` deduplicates the data by dropping the ordering of a given pair of sites (one of the two orderings) that yields a negative elevation difference while retaining the alternative ordering that yields a positive elevation difference. By >= 0, I retain comparisons of a site to itself. I drop (in the ggplot call) within-site comparisons when I plot against elevation difference, and I isolate within-site comparison when I plot against time. In the latter case, it is also necessary to deduplicate, since each pair of site dates occurs in A-B and B-A. For that, I filter (again, in the ggplot call) so that yday.diff > 0, similarly dropping the orderings that yield negative differences. Note that all this deduplication nonsense is not necessary for the betalinkr call because it deduplicates internally.

# Fun with betapart
fl_betapart <- survey %>%
  group_by(site, plant.sp) %>%
  summarize(count = sum(flower.cover)) %>%
  mutate(present = ifelse(count > 0, 1, 0)) %>%
  select(-count) %>%
  pivot_wider(names_from = plant.sp, values_from = present) %>%
  replace(is.na(.), 0) %>%
  column_to_rownames("site") %>%
  beta.pair(index.family = "jaccard")

fl_betapart.turnover <- fl_betapart$beta.jtu %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "tu") %>%
  filter(site1 != site2) %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, tu, tree_line1 = tree_line) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, tu, tree_line1, tree_line2 = tree_line)

fl_betapart.nestedness <- fl_betapart$beta.jne %>%
  as.matrix() %>%
  as.data.frame() %>%
  rownames_to_column(var = "site1") %>%
  pivot_longer(cols = -site1, names_to = "site2", values_to = "ne") %>%
  filter(site1 != site2) %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, ne, tree_line1 = tree_line) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, ne, tree_line1, tree_line2 = tree_line)

fl_betapart.collate <- full_join(fl_betapart.turnover, fl_betapart.nestedness) %>%
  mutate(tree_line.comp = case_when(
    tree_line1 == "below" & tree_line2 == "below" ~ "bb",
    tree_line1 == "above" & tree_line2 == "above" ~ "aa",
    tree_line1 != tree_line2 ~ "ab"
  )) %>%
  pivot_longer(cols = c(tu, ne), values_to = "value", names_to = "var")

ggplot(fl_betapart.collate, aes(var, value)) +
  geom_boxplot() +
  facet_wrap(~tree_line.comp)
```

## 3.2  Interaction beta-diversity and partitioning
S	= beta_S, the dissimilarity in species composition
OS = beta_OS, the dissimilarity (component) explained by "rewiring" among shared species
WN = beta_WN, the dissimilarity between the two networks
ST = beta_ST, the dissimilarity (component) explained by difference in species community composition 
```{r message=FALSE}
# Create webs
net_web <- web_prep(net)

@@

# Write to csv for archiving
#write_csv(beta_mit_data, "../Dryad_ms1/partitioned_beta_diversity")
```

# 4. Visualize beta-diversity ~ elevation diff
```{r}
# Set palette
p1 <- c("OS" = "#e7298a", "ST" = "#66a61e", "WN" = "black")
